### 2018-04-09 NC_NR_VC_VR/DESeq2/Interaction/makeDataMetrics-Test.Rmd

---
title: "Honey Bees - DESeq Interaction"
author: Lindsay Rutter
output:
  packagedocs::package_docs:
    toc: true
    toc_collapse: true
vignette: >
  %\VignetteEngine{packagedocs::redirect}
---

<meta http-equiv="content-type" content="text/html;charset=utf-8" />

```{r global_options, include=FALSE}
# R output pre blocks are styled by default to indicate output
knitr::opts_chunk$set(
  comment = NA,
  cache = TRUE,
  fig.height = 8,
  fig.width = 10
)
```

```{r}
library(DESeq2)
library(gplots)
library(RColorBrewer)
library(data.table)
library(ggplot2)

thisPath <- getwd()

data <- readRDS("../../data/data.Rds")
data <- as.matrix(data)

coldata = data.frame(row.names = colnames(data), virus = unlist(lapply(colnames(data), function (x) substring(unlist(strsplit(x, "[.]"))[1],1,1))), diet = unlist(lapply(colnames(data), function (x) substring(unlist(strsplit(x, "[.]"))[1],2,2))), treatment = unlist(lapply(colnames(data), function (x) unlist(strsplit(x, "[.]"))[1])))

dds = DESeqDataSetFromMatrix(countData = data, colData = coldata, design = ~ virus + diet + virus*diet)

dds <- DESeq(dds)

# Intercept, virus_V_vs_N, diet_R_vs_C, virusV.dietR
resultsNames(dds)

res <- results(dds, name="Intercept")
length(which((res@listData)$padj <0.05))

res <- results(dds, name="virus_V_vs_N")
length(which((res@listData)$padj <0.05))

res <- results(dds, name="diet_R_vs_C")
length(which((res@listData)$padj <0.05))

res <- results(dds, name="virusV.dietR")
length(which((res@listData)$padj <0.05))

```