### 2018-04-11 VirusHoneyBee/EdgeR/edgeR/makeDataMetrics.Rmd
Standard scatterplots Gailbraith data

---
title: "Honey Bees - edgeR"
author: Lindsay Rutter
output:
  packagedocs::package_docs:
    toc: true
    toc_collapse: true
vignette: >
  %\VignetteEngine{packagedocs::redirect}
---

<meta http-equiv="content-type" content="text/html;charset=utf-8" />

```{r global_options, include=FALSE}
# R output pre blocks are styled by default to indicate output
knitr::opts_chunk$set(
  comment = NA,
  cache = TRUE,
  fig.height = 8,
  fig.width = 10
)
```

```{r}
library(edgeR)
library(ggplot2)
library(GGally)
library(EDASeq)
library(utils)

thisPath <- getwd()

beeCounts <- readRDS("../../data/data.Rds")

beeCounts <- as.matrix(beeCounts)

y <- DGEList(counts=beeCounts)
```

edgeR vignette states that a gene is required to have a count of 5-10 in a library to be considered expressed in that library. Here minLib is equal to `r min(y$samples$lib.size)`. A CPM of 4 corresponds to a count of ~1 in the minimum number of samples in a group (6). So, I keep only rows that have that. This reduces the number of genes from 15,314 to 8,672.

```{r}
minLib <- min(y$samples$lib.size)
nRep = length(lapply(colnames(beeCounts), function (x) unlist(strsplit(x, "[.]"))[1])) / length(unique(lapply(colnames(beeCounts), function (x) unlist(strsplit(x, "[.]"))[1])))
keep <- rowSums(cpm(y)>round(minLib/1000000)) >= nRep/2
# Number of genes 15,314--> 8,672
y <- y[keep, , keep.lib.sizes=FALSE]
```

Next, I used edgeR normalization.
```{r}
y <- calcNormFactors(y)
```

I can create boxplot and RLE boxplot.

```{r}
ggparcoord(data.frame(log(y[[1]]/colMeans(y[[1]]))), columns=1:6, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

I can create MDS plots of interest.

```{r}
allGroups <- c(rep("C",3), rep("T",3))
y$samples$group <- allGroups

plotMDS(y, col = c("red","green","blue","purple")[factor(allGroups)], cex=0.6)
plotMDS(y, col = c("blue","blue","red","red")[factor(allGroups)], cex=0.6)
plotMDS(y, col = c("blue","red","blue","red")[factor(allGroups)], cex=0.6)
```

Copying Section (3.3.1 in edgeR vignette)

```{r}
Group = factor(c(rep("C",3), rep("T",3)))
#targets <- data.frame(rownames=paste0("Sample",1:24), Group = Group)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

```{r}
fit <- glmFit(y, design)
saveRDS(y[[1]], file="data.Rds")
```

Below we save the DEGs from all pairwise combinations of treatment groups.

```{r}
allPairs = data.frame(Treatment1 = factor(), Treatment2 = factor(), NumberDEG = numeric(), FirstLarger = numeric(), SecondLarger = numeric())

dataMetrics <- list()

for (i in 1:(ncol(fit)-1)){
  for (j in (i+1):ncol(fit)){
    contrast=rep(0,ncol(fit))
    contrast[i]=1
    contrast[j]=-1
    lrt <- glmLRT(fit, contrast=contrast)
    lrt <- topTags(lrt, n = nrow(y[[1]]))[[1]]
    lrt5 <- lrt[which(lrt$FDR<0.05),]

    df <- y[[1]][rownames(y[[1]]) %in% rownames(lrt5), c(which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[i])), which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[j])))]
    firstLarger <- length(which(apply(df, 1, function(x) mean(x[1:3]) > mean(x[4:6]))))
    secondLarger <- length(which(apply(df, 1, function(x) mean(x[1:3]) < mean(x[4:6]))))

    lrtLength <- nrow(lrt5)
    allPairs = rbind(allPairs, data.frame(Treatment1 = factor(colnames(fit)[i]), Treatment2 = factor(colnames(fit)[j]), NumberDEG = lrtLength, FirstLarger = firstLarger, SecondLarger = secondLarger))

    metrics = cbind(ID = rownames(lrt), lrt)
    metrics$ID = as.character(metrics$ID)
    dataMetrics[[paste0(colnames(fit)[i], "_", colnames(fit)[j])]] <- metrics
  }
}
allPairs <- allPairs[order(allPairs$NumberDEG),]
saveRDS(dataMetrics, file="dataMetrics.Rds")
```

```{r}
allPairs
```